---
project:
  title: "day-09"
  output-dir: docs
  
format:
  html:
    self-contained: true 
execute: 
  echo: true
---
```{r}
library(dplyr)
library(visdat)
library(tidyr)
library(ggplot2)
library(tidyverse)
```
```{r}
?airquality
vis_dat(airquality)
```
# The data does need cleaning.
```{r}
cleaned_airdata <- airquality %>% drop_na(Ozone) %>% drop_na(Solar.R)
vis_dat(cleaned_airdata)
summary(cleaned_airdata)
```
# The data is cleaned up now and I am choosing to use temperature as the predictor variable. I chose temperature because the ozone and temperature have a strong interaction as ozone formation is highly affected by temperature showing a positive correlation between the two.  
```{r}
model <- lm(Ozone ~ Temp, data = cleaned_airdata)
summary(model)
```
# I think this is a valid model since there was a meaningful R-squared value, significant p-value results and a reasonable residual behavior. 
# The R^2 means that about 48.8% of the variablility in Ozone levels is explained by the predictor of temperature. 
```{r}
a_data <- broom::augment(model, cleaned_airdata)
head(a_data)
```
```{r}
ggplot(a_data, aes(x = .fitted, y = Ozone)) +
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(title = "Actual VS Predicted Ozone Levels",
       subtitle = paste("Correlation:", round(cor(a_data$Ozone,
                                                  a_data$.fitted),2)),
       x = "Predicted Ozone Levels",
       y = "Actual Ozone Levels") +
  theme_minimal()
```
# The graph shows the linear relationship between the predicted and actual Ozone levels specifically showing a high correlation at 0.7. 
